Implementation of a stable, high-power optical lattice for quantum gas microscopy

### 2018-06-12 LatticeOptics/latticeOpticsv5/LatticeOpticsV5.tex
Fixed and updated references

Optical lattices holding an array of ultracold atoms have become a powerful experimental platform for a wide variety of investigations into various forms of quantum simulation~\cite{Friedenauer2008, Kim2010, Struck2011, Simon2011, Yan2013, Drewes2016, Murmann2015}.
The recent advent of quantum gas microscopy~\cite{Bakr2009, Sherson2010, Haller2015, Cheuk2015, Parsons2015, Edge2015, Omran2015, Greif2016, Cheuk2016, Parsons2016, Boll2016, Cheuk2016a, Brown2016} has enabled studies that exhibit unprecedented control over individual atoms at extremely low energy scales, particularly those governed by the Hubbard Hamiltonian~\cite{Hubbard1963}.
However, this progress has placed ever more stringent technical constraints on the lasers used to trap and manipulate the atomic systems.
For example, heating due to the intensity noise cannot be allowed to approach the longest thermalization timescales, lest the system fail to thermalize~\cite{Savard1997}.
Even if the system thermalizes, heating rates can limit the interrogation time available to probe a low-temperature state.
Similarly, experiments that rely on a small number of atoms held by multiple traps place stringent requirements on the relative positional stability of the traps, to ensure that the operations performed by the traps have adequate fidelity.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{Figure5.pdf}
    \caption{\textbf{Berek compensator based feedback loop}. This Berek compensator based feedback loop is used to vary the beam intensity at relatively high power. (a) Shows the small signal impulse response and (b-c) show the corresponding Bode plots.}\label{fig:berek_step_response}
  \end{center}
\end{figure}
A key point, as seen in Fig.~\ref{fig:berek}, is that if the waveplates are rotated together, the maximum transmitted power is fixed, while the minimum can be raised.
The waveplates are detuned such that the maximum power needed in \textit{interaction} mode is the minimum power transmitted by the Berek compensator system, strongly suppressing fluctuations due to angle positioning noise of the Berek compensator.
Naturally, rotation of a parallel plate in the beam path induces a shift of the beam, but this is acceptable because the Berek compensator is only active during the imaging phase of the experiment, when the position of the underlying harmonic trap is largely irrelevant.
Fortunately, the spatial phase of the optical lattice is set by the retro-reflector mirror, which does not move, and is not affected by the rotation of the Berek compensator.
This system functions as the actuator arm of a digital feedback loop, where the power of the beam is measured using a low noise photodiode. This feedback loop regulates the lattice between $1$ W of optical power and its maximum value (which depends on the amplifier, but is $\approx20$ W).
The step response and corresponding Bode plots~\cite{Bechhoefer2005} are shown in Fig.~\ref{fig:berek_step_response}.
A polarization rotation by $2 \pi$ is produced by approximately $3^\circ$ of axial rotation of the quartz plate.
Since we never need to rotate the plate outside this range, the position is constrained to this window in software.
Since two feedback loops are present in the system (the second one detailed in the next section), care must be taken so that they do not function simultaneously.
To accomplish this, the low power loop is always manually railed to its high extreme when the high power loop is active, and the digital high power loop is disabled when the low power loop is active.

@@

The connection between the actuation and monitoring mechanisms of the feedback system is the loop filter, which compares the \textit{set-point} signal, $v_s(t)$, supplied by the experimental control system and the monitoring signals $v_m(t)$, to adjust the actuation accordingly.
We use a simple, purely integrating design based in three stages, as seen in Fig.~\ref{fig:circuits}, based on TI's OPA211 op-amp.
The first stage computes and amplifies an error signal, $e(t)=v_m(t)-v_s(t)$, which is integrated by the second stage.
The integral gain is set using a low-drift trim-pot.
This is the case because within relevant frequency ranges, the AOM can be coarsely approximated by a pure time delay process (with delay given by the RF travel time across the crystal $\approx 1 \mu\mathrm{s}$).
In this scenario, the $P$ term must be small to avoid instability at the $\pi$ phase shift point, and a pure $I$ controller is optimal~\cite{Skogestad2001}.
For debugging purposes, the loop filter features buffered photodiode monitor and error signal monitor outputs.

>>>>

Optical lattices holding an array of ultracold atoms have become a powerful experimental platform for a wide variety of investigations into various forms of quantum simulation~\cite{Friedenauer2008, Kim2010, Struck2011, Simon2011, Yan2013, Drewes2016, Murmann2015}.
The recent advent of quantum gas microscopy~\cite{Bakr2009, Sherson2010, Haller2015, Cheuk2015, Parsons2015, Edge2015, Omran2015, Greif2016, Cheuk2016, Parsons2016, Boll2016, Cheuk2016a, Brown2017} has enabled studies that exhibit unprecedented control over individual atoms at extremely low energy scales, particularly those governed by the Hubbard Hamiltonian~\cite{Hubbard1963}.
However, this progress has placed ever more stringent technical constraints on the lasers used to trap and manipulate the atomic systems.
For example, heating due to the intensity noise cannot be allowed to approach the longest thermalization timescales, lest the system fail to thermalize~\cite{Savard1997}.
Even if the system thermalizes, heating rates can limit the interrogation time available to probe a low-temperature state.
Similarly, experiments that rely on a small number of atoms held by multiple traps place stringent requirements on the relative positional stability of the traps, to ensure that the operations performed by the traps have adequate fidelity.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{Figure5.pdf}
    \caption{\textbf{Berek compensator based feedback loop}. This Berek compensator based feedback loop is used to vary the beam intensity at relatively high power. (a) Shows the small signal impulse response and (b-c) show the corresponding Bode plots of the transfer function $G$.}\label{fig:berek_step_response}
  \end{center}
\end{figure}
A key point, as seen in Fig.~\ref{fig:berek}, is that if the waveplates are rotated together, the maximum transmitted power is fixed, while the minimum can be raised.
The waveplates are detuned such that the maximum power needed in \textit{interaction} mode is the minimum power transmitted by the Berek compensator system, strongly suppressing fluctuations due to angle positioning noise of the Berek compensator.
Naturally, rotation of a parallel plate in the beam path induces a shift of the beam, but this is acceptable because the Berek compensator is only active during the imaging phase of the experiment, when the position of the underlying harmonic trap is largely irrelevant.
Fortunately, the spatial phase of the optical lattice is set by the retro-reflector mirror, which does not move, and is not affected by the rotation of the Berek compensator.
This system functions as the actuator arm of a digital feedback loop, where the power of the beam is measured using a low noise photodiode. This feedback loop regulates the lattice between $1$ W of optical power and its maximum value (which depends on the amplifier, but is $\approx20$ W).
The step response and corresponding Bode plots~\cite{Bechhoefer2005} are shown in Fig.~\ref{fig:berek_step_response}.
A polarization rotation by $2 \pi$ is produced by approximately $3^\circ$ of axial rotation of the quartz plate.
Since we never need to rotate the plate outside this range, the position is constrained to this window in software.
Since two feedback loops are present in the system (the second one detailed in the next section), care must be taken so that they do not function simultaneously.
To accomplish this, the low power loop is always manually railed to its high extreme when the high power loop is active, and the digital high power loop is disabled when the low power loop is active.

@@

The connection between the actuation and monitoring mechanisms of the feedback system is the loop filter, which compares the \textit{set-point} signal, $v_s(t)$, supplied by the experimental control system and the monitoring signals $v_m(t)$, to adjust the actuation accordingly.
We use a simple, purely integrating design based in three stages, as seen in Fig.~\ref{fig:circuits}, based on TI's OPA211 op-amp.
The first stage computes and amplifies an error signal, $e(t)=v_m(t)-v_s(t)$, which is integrated by the second stage.
The integral gain is set using a low-drift trim-pot.
This is the case because within relevant frequency ranges, the AOM can be coarsely approximated by a pure time delay process (with delay given by the RF travel time across the crystal $\approx 1 \mu\mathrm{s}$).
In this scenario, the $P$ term must be small to avoid instability at the $\pi$ phase shift point, and a pure $I$ controller is optimal~\cite{Skogestad2003}.
For debugging purposes, the loop filter features buffered photodiode monitor and error signal monitor outputs.
